\documentclass[11pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{booktabs}
\usepackage{graphicx}
\graphicspath{{../figures/}}
\usepackage{xcolor}
\usepackage{subcaption}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{multirow}
\usepackage{hyperref}
\usepackage{cite}

\newtheorem{theorem}{Theorem}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{assumption}{Assumption}

% Custom commands
\newcommand{\Cs}{C_s}
\newcommand{\fmax}{f_{\max}}
\newcommand{\sstar}{s^{*}}
\newcommand{\pfault}{p_{\text{fault}}}
\newcommand{\rhoeff}{\rho_{\text{eff}}}
\newcommand{\tbase}{t_{\text{base}}}
\newcommand{\system}{\textsc{Prism}}

\begin{document}

\title{Capacity-Theoretic Limits of Slow Fault Mitigation\\under Power-of-Two-Choices Routing}
\author{DaLaw2}
\date{}
\maketitle

\begin{abstract}
Slow faults---where nodes remain reachable but exhibit degraded
performance---are a persistent source of tail latency violations in
distributed systems. Unlike crash faults, slow faults evade health checks
and produce ambiguous signals under high load. We present a
capacity-theoretic analysis of slow fault mitigation under
Power-of-Two-Choices (P2C) load balancing, establishing three results.
First, the \emph{operating envelope}: the maximum tolerable fault fraction
is $\fmax = 1 - L/U$, where $L$ is the load factor and $U$ the
utilization ceiling, validated empirically with zero fitted parameters.
Second, \emph{severity non-monotonicity}: under P2C routing, a moderate
slowdown (e.g., 10$\times$) causes worse tail latency than an extreme one
(e.g., 20$\times$), with the critical severity $\sstar =
fq/((1-q)(N-f))$. Third, \emph{SHED--P2C complementarity}: P2C's
queue-depth comparison and explicit weight reduction (SHED) are
complementary mechanisms operating at different stages of routing---SHED
provides 5--70\,ms P99 improvement at moderate severity ($2$--$5\times$)
but is unnecessary above $\sstar$ where P2C self-isolates faulty nodes.
Crucially, any detection-triggered weight reduction achieves nearly
identical results to full isolation under P2C, demonstrating that
\emph{detection speed, not response strategy, is the primary bottleneck}
in slow fault mitigation. Through discrete-event simulation across 15
fault scenarios, we show that reactive mechanisms (hedged requests, timeout
retry) are ineffective, while proactive detection with traffic exclusion is
necessary and sufficient. Our results provide practitioners with analytical
tools to determine when mitigation is feasible, which severity regime
demands intervention, and why detection precision matters less than
detection speed under queue-aware routing.
\end{abstract}

\input{sections/introduction}
\input{sections/background}
\input{sections/model}
\input{sections/analysis}
\input{sections/design}
\input{sections/evaluation}
\input{sections/related}
\input{sections/conclusion}

\bibliographystyle{plain}
\bibliography{references}

\end{document}
